Metabolomics

### 2021-04-28 inst/analysis/metabolomics.Rmd
differential metabolite analysis

```{r}
withr::with_dir(here::here(), {
  raw <- tar_read(metab_targeted_raw)
  clean <- tar_read(metab_targeted_clean)
})
```

@@

```{r}
pcaMethods::pca(t(SummarizedExperiment::assay(raw)), scale = "uv", center = TRUE) %>% 
  pcaMethods::scores() %>% 
  merge(SummarizedExperiment::colData(raw), by = 0) %>% 
  filter(type %nin% c("qc", "blank")) %>% 
  ggplot() +
  aes(
    x = PC1, 
    y = PC2, 
    shape = interaction(type, treatment, oxygen, sep = "|"),
    color = replicate
  ) + 
  geom_point()
```
## Cleaning

1. drift correction
1. remove low quality features (QC RSD > 0.2, dispersion ratio > 0.4)
1. impute missing values with random forest
1. probabilistic quotient normalization

```{r}
pcaMethods::pca(t(SummarizedExperiment::assay(clean)), scale = "uv", center = TRUE) %>% 
  pcaMethods::scores() %>% 
  merge(SummarizedExperiment::colData(clean), by = 0) %>% 
  ggplot() +
  aes(
    x = PC1, 
    y = PC2, 
    color = interaction(oxygen, treatment, sep = "|")
  ) + 
  geom_point() +
  labs(
    color = "Group"
  )
```

>>>>

```{r}
withr::with_dir(here::here(), {
  raw <- tar_read(metab_targeted_raw)
  clean <- tar_read(metab_targeted_clean)
  metab_pca <- tar_read(metab_targeted_pca)
  metab_limma <- tar_read(metab_targeted_limma)
  # metab_tt <- tar_read(metab_different_differences)
  volcano <- tar_read(metab_volcano)
})
```

@@

```{r}
pcaMethods::pca(t(SummarizedExperiment::assay(raw)), scale = "uv", center = TRUE) %>% 
  pcaMethods::scores() %>% 
  merge(SummarizedExperiment::colData(raw), by = 0) %>% 
  # filter(type %nin% c("qc", "blank")) %>% 
  ggplot() +
  aes(
    x = PC1, 
    y = PC2, 
    shape = type,
    color = interaction(oxygen, treatment, sep = "|")
  ) + 
  geom_point() +
  labs(
    color = "Group", 
    shape = "Type"
  )
```
## Cleaning

1. drift correction
1. remove low quality features (QC RSD > 0.2, dispersion ratio > 0.4)
1. impute missing values with random forest
1. probabilistic quotient normalization
1. batch correction

```{r}
metab_pca
```

## Differentially regulated metabolites

```{r}
volcano
```

```{r}
plot_mois(clean, "malate")
```
